---
title: "Poker Bot App Practice Lab"
author: "Draft for Mathematics of Poker"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE
)
```

# Poker Bot App Practice Lab

## Purpose

In this lab you will work through three connected ideas that sit at the heart of the poker bot app:

1. **Mathematical decision tools** such as pot odds, break-even equity, expected value, and regret.
2. **Ranges and uncertainty**, including Monte Carlo equity and weighted opponent ranges.
3. **Bot design**, where you inspect the information available to a bot and then modify a starter bot.

The goal is not to build a perfect poker agent. The goal is to understand how mathematical quantities become inputs to decisions, and how those decisions are encoded in an R function.

## Learning Goals

By the end of the lab, you should be able to:

- compute and interpret core poker quantities using the functions in the app,
- estimate hand equity using simulation,
- represent a simple weighted range in R,
- inspect the `bot_input` object passed to a bot,
- modify a simple bot so that its actions depend on mathematical information.

## Suggested Timing

A reasonable pacing for a 90-minute class period is:

- **Part I. Core calculations:** 20 minutes
- **Part II. Equity and ranges:** 25 minutes
- **Part III. Inspecting the bot input:** 15 minutes
- **Part IV. Modifying a bot:** 25 minutes
- **Wrap-up:** 5 minutes

---

# Part 0. Setup

Place this file inside the poker bot project folder. Then load the project files.

```{r}
source("poker_load_all.R")
poker_load_all(include_demos = TRUE)
```

If your project folder is set correctly, this should load the engine, math functions, example bots, and demos.

## Files you will use today

For reference, the most important files for this lab are:

- `poker_math.R`: core mathematical calculations
- `equity_tools.R`: Monte Carlo equity tools
- `quant_tools.R`: range utilities and additional quantitative tools
- `game_engine.R`: the tournament and hand state engine
- `example_bots.R`: sample bots
- `BotTemplate.R`: a student-facing template for writing a bot

---

# Part I. Core Poker Math

A bot does not directly "know" whether a play is good. Instead, it uses quantities that help evaluate decisions under uncertainty.

## 1. Pot odds and break-even equity

Suppose the pot is 120 chips and you must call 40 chips.

```{r}
pot_before_call <- 120
call_amount <- 40

pot_odds(call_amount, pot_before_call)
break_even_equity_call(call_amount, pot_before_call)
```

### Task 1

Interpret both outputs in complete sentences.

**Write here:**

- Pot odds measure:
- Break-even equity means:

## 2. Break-even fold probability for a bluff

Suppose the pot is 100 chips and you are considering a bluff of 75 chips.

```{r}
pot_before_bet <- 100
bet_amount <- 75

break_even_fold_prob_bluff(pot_before_bet, bet_amount)
```

### Task 2

Explain what this number means. Then decide whether the following statement is true:

> If your opponent folds more often than this threshold, the bluff is profitable in expectation.

**Write here:**

## 3. Expected value of a call or shove

Now suppose:

- the pot is 150,
- the amount to call is 50,
- your equity when called is 0.35.

```{r}
ev_call(equity = 0.35, call_amount = 50, pot_before_call = 150)
```

Try a few values of equity and see how the output changes.

```{r}
ev_call(equity = 0.20, call_amount = 50, pot_before_call = 150)
ev_call(equity = 0.30, call_amount = 50, pot_before_call = 150)
ev_call(equity = 0.40, call_amount = 50, pot_before_call = 150)
```

### Task 3

At about what equity does the call become profitable? Compare your answer to the break-even equity from Task 1.

**Write here:**

## 4. Regret as a way to compare actions

A bot may choose among several actions. One useful way to think is to compare the chosen action with the best available action.

```{r}
ev_options <- c(fold = 0, call = 8, raise = 13)
ev_best_action(ev_options)
ev_regret(ev_options, chosen_action = "call")
```

### Task 4

What is the regret of calling here? What does that regret mean in words?

**Write here:**

---

# Part II. Equity and Ranges

Poker decisions depend on hidden information. Since we do not know the opponent's exact hand, we work with **ranges** and with **equity** against possible holdings.

## 5. Monte Carlo hand-vs-hand equity

The project includes Monte Carlo equity tools for Texas Hold'em.

Here is a classic matchup: pocket aces against pocket kings.

```{r}
hole_1 <- data.frame(
  rank = c("A", "A"),
  suit = c("h", "s"),
  card = c("Ah", "As"),
  stringsAsFactors = FALSE
)

hole_2 <- data.frame(
  rank = c("K", "K"),
  suit = c("h", "s"),
  card = c("Kh", "Ks"),
  stringsAsFactors = FALSE
)

holdem_equity_mc_fast(
  hole_list = list(hole_1, hole_2),
  n_sims = 5000
)
```

### Task 5

Record the approximate equity of each hand. Then rerun the chunk with `n_sims = 500`, `5000`, and `20000`.

What changes as the number of simulations increases?

**Write here:**

## 6. Equity on a partial board

Now suppose the flop is already known.

```{r}
board_df <- data.frame(
  rank = c("A", "7", "2"),
  suit = c("d", "c", "h"),
  card = c("Ad", "7c", "2h"),
  stringsAsFactors = FALSE
)

holdem_equity_mc_fast(
  hole_list = list(hole_1, hole_2),
  board_df = board_df,
  n_sims = 5000
)
```

### Task 6

Why does Player 1's equity change so dramatically here? Use poker language if helpful, but be mathematically precise.

**Write here:**

## 7. Weighted ranges

A range is a collection of possible hands together with weights. In this project, a simple Hold'em range can be built with `new_range_holdem()`.

```{r}
example_range <- new_range_holdem(
  data.frame(
    c1 = c("Ah", "Ks", "Qh"),
    c2 = c("Kd", "Qc", "Qs"),
    w  = c(3, 2, 1)
  ),
  label = "Example weighted range"
)

example_range
example_range$combos
sum(example_range$weights)
```

Notice that the weights are normalized automatically.

### Task 7

Explain what it means that the weights are normalized. Why might weighted ranges be more realistic than treating every possible hand as equally likely?

**Write here:**

## 8. Build your own simple range

Create a range meant to represent a very strong preflop raising range. Use 4 to 6 specific two-card combinations and give larger weights to the strongest hands.

```{r}
strong_range <- new_range_holdem(
  data.frame(
    c1 = c("Ah", "As", "Kh", "Ad"),
    c2 = c("Ac", "Kd", "Ks", "Kc"),
    w  = c(4, 4, 3, 2)
  ),
  label = "Strong opening range"
)

strong_range$combos
```

### Task 8

Modify the example above to create:

- a **tight range**, and
- a **loose range**.

Then compare the size and the weight distribution of the two.

```{r}
# Replace this with your own construction.
tight_range <- strong_range
loose_range <- strong_range

range_size(tight_range)
range_size(loose_range)

tight_range$combos
loose_range$combos
```

**Write here:**

## 9. Board texture features

The project also includes simple board-texture tools.

```{r}
flop_df <- data.frame(
  rank = c("K", "T", "4"),
  suit = c("h", "h", "c"),
  card = c("Kh", "Th", "4c"),
  stringsAsFactors = FALSE
)

board_features(flop_df)
```

### Task 9

Inspect the output. Which pieces of this output might be useful to a bot deciding whether to bet the flop?

**Write here:**

---

# Part III. What Information Does a Bot Actually Receive?

A poker bot does not see the whole tournament state directly. It is given a structured object called `bot_input`.

## 10. Create a live tournament state

We will initialize a small tournament, start a hand, and inspect the current acting player's input.

```{r}
bot_fns <- list(
  "Random Bot" = random_bot,
  "Caller Bot" = always_call_bot,
  "Passive Bot" = passive_bot
)

tourn <- initialize_tournament(
  bot_fns = bot_fns,
  player_names = names(bot_fns),
  starting_stack = 1000
)

tourn <- initialize_hand(tourn)
tourn <- post_blinds_and_antes(tourn)
```

Now build the input that the acting bot will receive.

```{r}
bot_input_example <- build_bot_input(tourn)
str(bot_input_example, max.level = 2)
```

You can also view it as a data frame.

```{r}
demo_show_bot_input(tourn)
```

### Task 10

Find and record the following pieces of information inside `bot_input_example`:

- your hole cards,
- the current pot,
- the current street,
- the legal actions,
- your current stack,
- the public information about the other players.

**Write here:**

## 11. Explore the legal action structure

```{r}
bot_input_example$legal_actions
bot_input_example$legal_actions$legal_action_types
bot_input_example$legal_actions$actions
```

### Task 11

Why is it important for a bot to check which actions are legal before returning an action?

**Write here:**

---

# Part IV. Reading and Modifying a Bot

## 12. Read a starter bot

Open `example_bots.R` and locate the function `simple_preflop_strength_bot()`.

This bot does something simple:

- if it is preflop and it likes its hand, it plays aggressively,
- otherwise it falls back to checking or calling when possible.

### Task 12

Read the function and answer:

1. What counts as a "premium" hand in this bot?
2. What does the bot do with premium hands?
3. What does it do after the flop?

**Write here:**

## 13. Try a starter template

Open `BotTemplate.R`. You will see a function called `my_bot_name()` and helper functions such as:

- `bot_has_action()`
- `bot_min_bet()`
- `bot_min_raise()`
- `choose_preferred_action()`

These help keep your bot legal and readable.

## 14. Build a simple math-based bot

Create a new bot that follows this rule:

- **Preflop:**
  - raise minimum with pairs, ace-king, or ace-queen,
  - otherwise check if possible,
  - otherwise call if the call is at most one big blind,
  - otherwise fold.
- **Postflop:**
  - check when possible,
  - otherwise call only if the break-even equity threshold is at most `0.25`,
  - otherwise fold.

Below is a starter version. Complete or revise it.

```{r}
lab_bot <- function(bot_input) {
  legal_types <- bot_input$legal_actions$legal_action_types
  hole_cards <- bot_input$hole_cards
  street <- bot_input$street
  big_blind <- bot_input$big_blind
  pot <- bot_input$pot
  current_bet <- bot_input$current_bet
  committed <- bot_input$committed_this_round

  call_amount <- max(0, current_bet - committed)
  vals <- sort(hole_rank_values(hole_cards), decreasing = TRUE)

  if (street == "preflop" && length(vals) == 2) {
    paired <- vals[1] == vals[2]
    ak <- identical(vals, c(14, 13))
    aq <- identical(vals, c(14, 12))

    if (paired || ak || aq) {
      if (bot_has_action(bot_input, "raise")) {
        return(list(type = "raise", amount = bot_min_raise(bot_input)))
      }
      if (bot_has_action(bot_input, "bet")) {
        return(list(type = "bet", amount = bot_min_bet(bot_input)))
      }
    }

    if ("check" %in% legal_types) {
      return(list(type = "check"))
    }

    if ("call" %in% legal_types && call_amount <= big_blind) {
      return(list(type = "call"))
    }

    return(list(type = "fold"))
  }

  if ("check" %in% legal_types) {
    return(list(type = "check"))
  }

  if ("call" %in% legal_types) {
    threshold <- break_even_equity_call(call_amount, pot)
    if (threshold <= 0.25) {
      return(list(type = "call"))
    }
  }

  list(type = "fold")
}
```

### Task 13

Before running the bot, explain the logic of the line

```r
threshold <- break_even_equity_call(call_amount, pot)
```

What is the bot using this threshold for?

**Write here:**

## 15. Test your bot on a live input

Replace one of the bots in the tournament with your new bot.

```{r}
bot_fns_test <- list(
  "Lab Bot" = lab_bot,
  "Random Bot" = random_bot,
  "Caller Bot" = always_call_bot
)

tourn2 <- initialize_tournament(
  bot_fns = bot_fns_test,
  player_names = names(bot_fns_test),
  starting_stack = 1000
)

tourn2 <- initialize_hand(tourn2)
tourn2 <- post_blinds_and_antes(tourn2)

bot_input_test <- build_bot_input(tourn2)
demo_show_bot_input(tourn2)
lab_bot(bot_input_test)
```

### Task 14

Run the chunk several times by re-initializing the tournament. Does your bot always return a legal action? Describe one situation where your bot raises, one where it calls, and one where it folds.

**Write here:**

## 16. Optional challenge: make the bot more thoughtful

Revise `lab_bot()` so that it also uses **board texture** on the flop.

For example, you might decide that on the flop the bot should:

- bet when checked to on dry boards,
- check more often on coordinated two-tone boards,
- call less often when the required equity threshold is large.

A starter idea is below.

```{r}
lab_bot_v2 <- function(bot_input) {
  legal_types <- bot_input$legal_actions$legal_action_types
  board <- bot_input$board
  street <- bot_input$street
  pot <- bot_input$pot
  current_bet <- bot_input$current_bet
  committed <- bot_input$committed_this_round
  call_amount <- max(0, current_bet - committed)

  if (street == "flop" && length(board) == 3) {
    board_df <- parse_cards(board)
    feats <- board_features(board_df)

    if ("bet" %in% legal_types && !isTRUE(feats$two_tone) && feats$connectivity <= 1) {
      return(list(type = "bet", amount = bot_min_bet(bot_input)))
    }
  }

  if ("check" %in% legal_types) return(list(type = "check"))

  if ("call" %in% legal_types) {
    threshold <- break_even_equity_call(call_amount, pot)
    if (threshold <= 0.20) return(list(type = "call"))
  }

  list(type = "fold")
}
```

### Task 15

Modify this bot and explain your design choices.

**Write here:**

---

# Wrap-Up Questions

Answer the following in a few sentences each.

## 17. Math to code

Which part of today's lab felt most clearly mathematical, and which part felt most like programming? How did the two interact?

**Write here:**

## 18. Limits of the current bots

What is one important piece of information that your bot does **not** currently use, but probably should use in a more serious version?

**Write here:**

## 19. Reflection

A poker bot acts under uncertainty and with limited information. Give one example from today's lab where the bot had to rely on a model or approximation rather than exact knowledge.

**Write here:**

---

# Submission Option

If your instructor asks for a submission, turn in:

1. this completed lab file, and
2. one revised bot function with a short paragraph explaining its logic.

# Instructor Notes (optional to keep or delete)

This lab was written to match the current app structure, especially the following functions and files:

- `break_even_equity_call()`, `ev_call()`, `ev_regret()` from `poker_math.R` and `quant_tools.R`
- `holdem_equity_mc_fast()` from `equity_tools.R`
- `new_range_holdem()` and `board_features()` from `quant_tools.R`
- `build_bot_input()`, `bot_input_to_dataframe()`, and `demo_show_bot_input()` from `game_engine.R`
- `simple_preflop_strength_bot()`, `random_bot()`, `always_call_bot()`, and `passive_bot()` from `example_bots.R`
- the student-facing structure in `BotTemplate.R`

